\textbf{Proof Reconstruction in Classical Propositional Logic}

### 2017-05-09 slides/slides.tex
[ slides ] last corrections.

% footer.
\setbeamercolor{headFoot}{fg=white, bg=plum!80!black}
\setbeamertemplate{footline}{
  \leavevmode%
  \hbox{%
  \begin{beamercolorbox}
    [wd=.8\paperwidth,ht=2.3ex,dp=1ex,left]{headFoot}%
    \hspace*{2ex}\textbf\insertshorttitle\hspace*{2mm}|\hspace*{2mm}\textbf\insertshortauthor
  \end{beamercolorbox}%
  \begin{beamercolorbox}
    [wd=.2\paperwidth,ht=2.3ex,dp=1ex,right]{headFoot}%
    \insertframenumber{}/\inserttotalframenumber\hspace*{2ex}
  \end{beamercolorbox}}%
  \vskip 0pt%
}

@@

\setbeamerfont{framesubtitle}{size=\normalfont\tiny\itshape}
\setbeamercolor{framesubtitle}{fg=white, bg=plum}

@@

\texttt{Metis} is an automatic theorem prover for First-Order Logic with Equality
\citep{hurd2003first}
\begin{block}{Why Metis?}
  \begin{itemize}
    \item \href{https://github.com/gilith/metis}{Open source} implemented in
       \texttt{Standard ML}
    \item Each refutation step is one of
      \href{http://www.gilith.com/software/metis/logical-kernel.txt}{\textit{6 rules}}
    \item Reads problem in \hyperlink{tptp-syntax}{\texttt{TPTP} format}
    \item Outputs detailed proofs in \hyperlink{tstp-syntax}{\texttt{TSTP} format}
  \end{itemize}
\end{block}
\gobackreconstruction
\end{frame}

@@

\begin{frame}[fragile, label=atp-conjunct]{\texttt{Agda-Metis}: Conjunct Inference
  \footnote{\url{https://github.com/jonaprieto/agda-metis}}}
\begin{itemize}
\item Definition
  \begin{equation*}
  conjunct (ϕ₁ ∧  ϕ₂ ∧ ⋯ ∧ \mathbf{ϕᵢ} ∧ ⋯ ∧ \phi_n, \mathbf{ϕᵢ}) ⟶ \mathbf{ϕᵢ}
  \end{equation*}

@@

\begin{frame}[fragile, label=verified-example-4]{$p ∧ q ⊢ q ∧ p$}{Reconstructed proof}
\begin{minted}{cagda}
Γ : Ctxt
Γ = [ a ]

@@

\begin{frame}[fragile, label=failure-example]{Bug in the Printing of the Proof}
  {\texttt{Metis}' Issue: \url{https://github.com/gilith/metis/issues/2}}
% {}
% \begin{tptp}
% $ cat problem.tptp
% fof(goal, conjecture,
%   ((p <=> q) <=> r) <=> (p <=> (q <=> r))).
% \end{tptp}
\begin{tptp}
$ metis --version
metis 2.3 (release 20161108)
$ metis --show proof problem.tptp
...
fof(normalize_2_0, plain,
  (~ p & (~ q <=> ~ r) & (~ p <=> (~ q <=> ~ r))),
  inference(canonicalize, [], [negate_2_0])).
fof(normalize_2_1, plain, ~ p <=> (~ q <=> ~ r),
    inference(conjunct, [], [normalize_2_0])).
fof(normalize_2_2, plain, ~ q <=> ~ r,
    inference(conjunct, [], [normalize_2_0])).
fof(normalize_2_3, plain, ~ p,
  inference(conjunct, [], [normalize_2_0])).
fof(normalize_2_4, plain, $false,
    inference(simplify, [],
      [normalize_2_1, normalize_2_2, normalize_2_3])).
...
\end{tptp}
\end{frame}

@@

\only<5->{ \node[below = 0.8cm of apia] (atp)
  {\includegraphics[scale=0.8]{figures/atp}}};

@@

\only<2->{\draw[->, ultra thick] (agdafile) to (eagda)};
\only<2->{\draw[->, ultra thick] (eagda) to (agdai)};
\only<3->{\draw[->, ultra thick] (agdai) to (apia)};
\only<4->{\draw[->, very thick, dashed, gray] (apia) to (tptp)};
\only<5->{\draw[->, very thick, dashed, gray] (tptp) to (atp)};
\only<6->{\draw[->, very thick, dashed, gray] (atp) to (apia)};

\end{tikzpicture}
\end{frame}

@@

\begin{frame}[label=future-work]{Future Work}
\begin{itemize}
  \item Integration with \texttt{Apia}
  \item Support First-Order Logic with Equality
  \item Support another prover like \texttt{EProver} or \texttt{Vampire}
\end{itemize}
\end{frame}

\begin{frame}
\end{frame}

\begin{frame}[label=references, allowframebreaks]{References}
\printbibliography
\end{frame}

\end{document}

>>>>

% footer.
\setbeamercolor{headFoot}{fg=white, bg=plum!80!black}
\setbeamertemplate{footline}{
  \leavevmode%
  \hbox{%
  \begin{beamercolorbox}
    [wd=.8\paperwidth,ht=2.3ex,dp=1ex,left]{headFoot}%
    \hspace*{2ex}\textbf\insertshorttitle\hspace*{2mm}|
    \hspace*{2mm}\textbf\insertshortauthor
  \end{beamercolorbox}%
  \begin{beamercolorbox}
    [wd=.2\paperwidth,ht=2.3ex,dp=1ex,right]{headFoot}%
    \insertframenumber{}/\inserttotalframenumber\hspace*{2ex}
  \end{beamercolorbox}}%
  \vskip 0pt%
}

@@

\setbeamerfont{framesubtitle}{size=\normalfont\tiny}
\setbeamercolor{framesubtitle}{fg=white, bg=plum!80!black}

@@

\texttt{Metis} is an automatic theorem prover for First-Order Logic with Equality
\citep{hurd2003first}
\begin{block}{Why Metis?}
  \begin{itemize}
    \item \href{https://github.com/gilith/metis}{Open source} implemented in
       \texttt{Standard ML}
    \item Each refutation step is one of
      \hyperlink{metis-inference-rules}{\textit{six rules}}
    \item Reads problem in \hyperlink{tptp-syntax}{\texttt{TPTP} format}
    \item Outputs detailed proofs in \hyperlink{tstp-syntax}{\texttt{TSTP} format}
  \end{itemize}
\end{block}
\gobackreconstruction
\end{frame}

@@

\begin{frame}[fragile, label=atp-conjunct]{\texttt{Agda-Metis}: Conjunct Inference
  \footnote{\url{https://github.com/jonaprieto/agda-metis}}}
\begin{itemize}
\item Definition
  \begin{equation*}
  \text{conjunct}(ϕ₁ ∧  ϕ₂ ∧ ⋯ ∧ \mathbf{ϕᵢ} ∧ ⋯ ∧ \phi_n, \mathbf{ϕᵢ}) ⟶ \mathbf{ϕᵢ}
  \end{equation*}

@@

\begin{frame}[fragile, label=verified-example-4]{$p ∧ q ⊢ q ∧ p$}{Reconstructed proof}
\begin{minted}{cagda}
-- Premise.

subgoal₀ = q
subgoal₁ = (q ⇒ p)

-- Proof

@@

\begin{frame}[fragile, label=failure-example]{Bug in the Printing of the Proof}
  {\texttt{Metis}' v2.3 (release 20161108) - Issue\footnote{\url{https://github.com/gilith/metis/issues/2}}}
% {}
\begin{tptp}
$ cat problem.tptp
fof(goal, conjecture,
  ((p <=> q) <=> r) <=> (p <=> (q <=> r))).
\end{tptp}
\begin{tptp}
$ metis --show proof problem.tptp
...
fof(normalize_2_0, plain,
  (~ p & (~ q <=> ~ r) & (~ p <=> (~ q <=> ~ r))),
  inference(canonicalize, [], [negate_2_0])).
fof(normalize_2_1, plain, ~ p <=> (~ q <=> ~ r),
    inference(conjunct, [], [normalize_2_0])).
fof(normalize_2_2, plain, ~ q <=> ~ r,
    inference(conjunct, [], [normalize_2_0])).
fof(normalize_2_3, plain, ~ p,
  inference(conjunct, [], [normalize_2_0])).
fof(normalize_2_4, plain, $false,
    inference(simplify, [],
      [normalize_2_1, normalize_2_2, normalize_2_3])).
...
\end{tptp}
\end{frame}

\begin{frame}[fragile, label=failure-example]{Bug in the Printing of the Proof}
  {\texttt{Metis}' v2.3 (release 20161108) - Issue\footnote{\url{https://github.com/gilith/metis/issues/2}}}

\[ φ := ¬ p ∧ (¬ q ⇔ ¬ r) ∧ (¬ p ⇔ (¬ q ⇔ ¬ r))\]

\begin{prooftree}
\AxiomC{$\vdots$}
\RightLabel{\scriptsize{canonicalize}}
\UnaryInfC{$φ$}
\RightLabel{\scriptsize{conjunct}}
\UnaryInfC{$¬ p ⇔ (¬ q ⇔ ¬ r)$}

\AxiomC{$\vdots$}
\RightLabel{\scriptsize{canonicalize}}
\UnaryInfC{$φ$}
\RightLabel{\scriptsize{conjunct}}
\UnaryInfC{$¬ q ⇔ ¬ r$}

\AxiomC{$\vdots$}
\RightLabel{\scriptsize{canonicalize}}
\UnaryInfC{$φ$}
\RightLabel{\scriptsize{conjunct}}
\UnaryInfC{$¬ p$}

\RightLabel{\scriptsize{simplify}}
\TrinaryInfC{$\bot$}
\end{prooftree}
\pause
The bug was caused by the conversion of \texttt{Xor} sets to \texttt{Iff} lists.
After reporting this, Joe fixed the printing of \texttt{canonicalize} inference rule
\[ φ := ¬ p ∧ (¬ q ⇔ ¬ r) ∧ (¬ p ⇔ (¬ q ⇔ {\color{red} r}))\]

\end{frame}

@@

\only<5->{ \node[below = 0.8cm of apia] (atp)
  {\includegraphics[scale=0.8]{figures/atp}
  \footnote{\url{http://github.com/jonaprieto/online-atps}}}
  };

@@

\only<2->{\draw[->, ultra thick] (agdafile) to (eagda)};
\only<2->{\draw[->, ultra thick] (eagda) to (agdai)};
\only<3->{\draw[->, ultra thick] (agdai) to (apia)};
\only<4->{\draw[->, very thick, dashed, gray] (apia) to (tptp)};
\only<5->{\draw[->, very thick, dashed, gray] (tptp) to (atp)};
\only<6->{\draw[->, very thick, dashed, gray] (atp) to (apia)};
\end{tikzpicture}
\end{frame}

@@

\begin{frame}[label=future-work]{Contributions and Future Work}
\vfill
\begin{table}[!ht]
\begin{center}
\scalebox{0.8}{
{\renewcommand{\arraystretch}{1.3}%
\begin{tabular}{|ll|}
\hline
\textbf{Name} & \textbf{Purpose} \\
\hline
\href{http://github.com/jonaprieto/agda-prop}{\texttt{Agda-Prop}}     & lib. for syntax and theorems of Clasical Propositional Logic\\
\href{http://github.com/jonaprieto/agda-metis}{\texttt{Agda-Metis}}   & lib. for versions of the inference rules used by \texttt{Metis}\\
\href{http://github.com/jonaprieto/athena}{\texttt{Athena}}           & soft. Translator in Haskell for \texttt{Metis'} \texttt{TSTP} files to \texttt{Agda}\\
\href{http://github.com/jonaprieto/online-atps}{\texttt{OnlineATPs}}  & soft. A client in Haskell for \texttt{SystemOnTPTP} of \texttt{TPTP} World\\[1ex]
\hline
\end{tabular}}
}
\end{center}
\label{tab:library}
\end{table}
\vfill

\begin{itemize}
  \item Integration with \texttt{Apia}
  \item Support First-Order Logic with Equality
  \item Support another prover like \texttt{EProver} or \texttt{Vampire}
\end{itemize}
\end{frame}

\begin{frame}[fragile, label=references, allowframebreaks]{References}
\printbibliography
\end{frame}

\begin{frame}[fragile,label=metis-inference-rules]{\texttt{Metis}' Inference Rules}
\vfill
\begin{prooftree}
\AxiomC{}
\RightLabel{axiom}
\UnaryInfC{$C$}
\end{prooftree}
\vfill
\begin{prooftree}
\AxiomC{}
\RightLabel{assume $L$}
\UnaryInfC{$L ∨ ¬ L$}
\end{prooftree}
\vfill
\begin{prooftree}
\AxiomC{$C$}
\RightLabel{subst $σ$}
\UnaryInfC{$σC$}
\end{prooftree}
\vfill
\begin{prooftree}
\AxiomC{$L ∨ C$}
\AxiomC{$¬ L ∨ D$}
\RightLabel{resolve $L$}
\BinaryInfC{$C ∨ D$}
\end{prooftree}
\vfill
\begin{prooftree}
\AxiomC{}
\RightLabel{refl $t$}
\UnaryInfC{$t = t$}
\end{prooftree}
\vfill
\begin{prooftree}
\AxiomC{}
\RightLabel{eq $L$ $p$ $t$}
\UnaryInfC{$¬ (L[p] = t) ∨ ¬ L ∨ L[ p ↦ t]$}
\end{prooftree}
\vfill
\end{frame}
 \begin{tikzpicture}[overlay, remember picture, scale=0.5]
    \tikzset{shift={(current page.center)}, xshift=11.8cm,yshift=-8.4cm}
    \node[fill=plum,text=white, hyperlink node = metis]
      (go) at (0,0) {\tiny Go Back};
  \end{tikzpicture}
\end{document}
